Assessing environmental and predation impacts on Antarctic Krill (*Euphausia superba*) dynamic population from an integrated Catch-at-Length model perspective

### 2025-04-03 MS2_Krill.Rmd
adding text

---
title: "Assessing environmental and predation effects on Antarctic Krill (*Euphausia superba*) stock productivity: An Integrated Catch-at-Length Model Perspective"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Erica Jarvis Mason^[Antarctic División NOAA, USA]
  - Andrea Pinones ^[Instituto Antártico Chileno, Instituto Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base]
output:
  bookdown::word_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
#csl: apa.csl
csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

```{r setup1, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(eval=FALSE,
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

```{r message=FALSE, eval=TRUE}
library(here)
library(kableExtra)
library(ggpubr)
library(tibble)
library(readxl)
library(openxlsx)
library(r4ss)
```

@@

Understanding the drivers of krill (*Euphausia superba*) population dynamics is crucial for sustainable management. This study evaluates the impact of environmental variables and predation components on krill productivity and other population parameters, in the Western Antarctic Peninsula (WAP). Using an integrated Catch-at-Length model, we assess how ecosystem factors influence recruitment, biomass trends, and stock status.  Our results indicate that krill population variables, as well as productivity estimates, tend to be more optimistic in terms of stock status when environmental and predators variables are not considered. This highlights a critical challenge in developing management recommendations and harvest control rules, as ignoring ecosystem influences may lead to overly optimistic assessments of krill stock health.  Our findings emphasize the necessity of incorporating ecosystem variables into population stock assessments process, as neglecting these factors may lead to biased estimations of krill population dynamics. We consider that the use of integrated models, which leverage the extensive population, fisheries, and ecosystem data available through various scientific programs, can provide valuable insights for management aligned with CCAMLR conservation and harvesting objectives.

*Keywords: Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, Catch-to-Length model, environmental factors, predation components, fisheries management.*

@@

There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture spatial variability, and account for ecosystem interactions. However, integrating these complexities presents several challenges. Long-term data consistency is often lacking due to variations in monitoring programs and sampling methodologies. Additionally, some areas have sparse or incomplete datasets on krill dynamics. Furthermore, indirect drivers, such as environmental variability and predator abundance, introduce another layer of complexity. Despite these gaps, a substantial amount of data exists on krill populations and their ecosystem, making it crucial to identify knowledge gaps and integrate available information into comprehensive models that support management recommendations. This requires not only refining existing datasets but also adopting a fully ecosystem-based management approach.

To address this need, we propose a conceptual framework for developing a statistical model that integrates multiple data sources to assess the population dynamics of Antarctic krill in the Western Antarctic Peninsula (WAP) over the past four decades. This approach incorporates biological, fishery, survey, and ecosystem-related information to provide a more comprehensive understanding of krill stock fluctuations. Specifically, we employ an integrated catch-at-length stock synthesis model that allows us to evaluate the effects of environmental drivers on krill recruitment and productivity, as well as the impact of predator-prey interactions on overall population dynamics.  Our approach involves comparing different scenarios, one incorporating ecosystem-related variables and another based solely on traditional stock assessment data, to determine how environmental fluctuations and predator abundance influence key population variables and parameters. By analyzing differences in estimated biomass, recruitment patterns, and overall krill productivity across these scenarios, we aim to quantify the extent to which ecosystem factors drive krill population trends. This comparative approach enables us to assess the relative importance of environmental conditions and predator dynamics in shaping krill stocks, providing critical insights for both fisheries management and conservation efforts.

By integrating ecosystem drivers into the stock assessment process, we enhance the accuracy of population estimates and improve our ability to predict future stock trends under changing environmental conditions. This broader ecological perspective strengthens the scientific foundation for sustainable management recommendations, ensuring that harvesting strategies align with an ecosystem-based management approach. Ultimately, the insights gained from this integrated modeling effort contribute to the ongoing efforts of CCAMLR's new strategy to balance krill fishery sustainability with the conservation of the broader Antarctic marine ecosystem.

@@

The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling. All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl. Mat. 2); 3) For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP. We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl. Mat 3); 4) Predator abundance data from "MAPPPD". Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].

A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously. Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population (Figure \@ref(fig:concem)).Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.

@@

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables. The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21. SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions. SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population. All analysis were executed in  R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023]. Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit). Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).

We applied an integrated Catch-to-Length model to estimate krill stock-recruitment relationships and population trends. The model links catch-at-length data with growth, mortality, and recruitment processes while incorporating environmental and predation variables. Model performance was evaluated through likelihood-based comparisons, sensitivity analyses, and cross-validation.   
Key model parameters include:

### Natural Mortality and Growth Parameterization

The natural mortality rate (\(M\)) was specified using a single parameter per morphotype. Growth was modeled following a von Bertalanffy growth function with size-at-age parameterization. The model assumed linear growth below a predefined settlement age of 1.5 years. The asymptotic length (\(L_{\infty}\)) was determined by the estimated length-at-age of 999 years, indicating that the model used \(L_{\infty}\) as an upper growth boundary. Exponential decay of growth beyond maximum age was disabled (, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints.

Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.

@@

Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years. Fecundity was estimated as a power function of body weight, where fecundity was computed as \(Eggs = Wt \times (a + b \times Wt)\). The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan.

The model used a direct parameterization approach for natural mortality (\(M\)), growth (\(G\)), and growth variability (\(CV_G\)), without additional offsets between sexes. Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters.

@@

Selectivity functions were modeled using a length-based logistic formulation for the fishery, survey, and predator fleets. The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width).

The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width. The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm. Parameter estimation phases were assigned, with negative values indicating fixed parameters.

The survey selectivity was also modeled using a length-based logistic function. Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm. The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values).

The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm. These parameters were estimated in later phases (positive phase values), allowing flexibility in their estimation.

Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization.

@@

The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation (option 3). This model was selected due to its robustness in representing recruitment dynamics in response to spawning biomass.

The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value ($SR\_LN(R_0)$) of 23, with a prior of 10. The parameter was freely estimated in phase 1. The steepness of the stock-recruitment curve ($SR\_BH\_steep$) was fixed at 0.85 in phase -4, ensuring consistency with previous studies. The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability. The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts. The $SR\_autocorr$ parameter was set to 0 and fixed, indicating that no temporal autocorrelation in recruitment deviations was assumed.

Recruitment deviations were estimated to account for interannual variability in recruitment dynamics. The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term. The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years.

The recruitment deviation phase was set to 2. Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed. Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment. The maximum bias adjustment factor was 0.8274.

To prevent overparameterization, recruitment deviations were constrained between -5 and 5. No cyclic patterns in recruitment were assumed, and advanced recruitment options were enabled for model flexibility.

@@

### Stock Assessment Scenarios

>>>>

---
title: "Assessing environmental and predation impacts on Antarctic Krill (*Euphausia superba*) dynamic population from an integrated Catch-at-Length model perspective"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Erica Jarvis Mason^[Antarctic División NOAA, USA]
  - Andrea Pinones ^[Instituto Antártico Chileno, Instituto Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base]
output:
  bookdown::word_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
#csl: apa.csl
csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

```{r setup1, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(eval=FALSE,
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      fig.show='hold',
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
```

```{r message=FALSE, eval=TRUE}
library(here)
library(kableExtra)
library(ggpubr)
library(tibble)
library(readxl)
library(openxlsx)
library(r4ss)
library(readr)
library(flextable)
library(officer)
library(r4ss)
```

@@

Understanding the drivers of krill (*Euphausia superba*) population dynamics is crucial for sustainable management. This study evaluates the impact of environmental variables and predation components on krill productivity and other population parameters, in the Western Antarctic Peninsula (WAP). Using an integrated Catch-at-Length model, we assess how ecosystem factors influence recruitment, biomass trends, and stock status. Our results indicate that krill population variables, as well as productivity estimates, tend to be more optimistic in terms of stock status when environmental and predators variables are not considered. Our findings emphasize the necessity of incorporating ecosystem variables into population stock assessments process, as neglecting these factors may lead to biased estimations of krill population dynamics. This highlights a critical challenge in developing management recommendations and harvest control rules, as ignoring ecosystem influences may lead to overly optimistic assessments of krill stock health. We consider that the use of integrated models, which leverage the extensive population, fisheries, and ecosystem data available through various scientific programs, can provide valuable insights for management aligned with CCAMLR conservation and harvesting objectives.

*Keywords: Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, SS3, Catch-to-Length model, environmental factors, predation components, fisheries management.*

@@

There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions. However, integrating these complexities presents several challenges. Long-term data consistency is often lacking due to variations in monitoring programs and sampling methodologies. Additionally, some areas have sparse or incomplete datasets on krill dynamics. Furthermore, indirect drivers, such as environmental variability and predator abundance, introduce another layer of complexity. Despite these gaps, a substantial amount of data exists on krill populations and their ecosystem, making it crucial to identify knowledge gaps and integrate available information into comprehensive models that support management recommendations. This requires not only refining existing datasets but also adopting a fully ecosystem-based management approach.

To address this need, we propose a conceptual framework for developing a statistical model that integrates multiple data sources to assess the population dynamics of Antarctic krill in the Western Antarctic Peninsula (WAP) over the past decades. This approach incorporates biological, fishery, survey, and ecosystem-related information to provide a more comprehensive understanding of krill stock fluctuations. Specifically, we employ an integrated catch-at-length implemented in Stock Synthesis model [@Methot2013] that allows us to evaluate the effects of environmental drivers on recruitment and productivity, as well as the impact of predator components into krill population dynamics. The primary objective of this study was to assess the statistical performance of different management metrics, population variables and parameters based on alternative approaches to incorporating both environmental and predator data into a typical krill stock assessment model. By analyzing differences in estimated biomass, recruitment patterns, and overall krill productivity across these scenarios, we aim to quantify the extent to which ecosystem factors drive krill population trends. This comparative approach enables us to assess the relative importance of environmental conditions and predator dynamics in shaping krill stocks, providing critical insights for both fisheries management and conservation efforts.

By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate conditions. This broader ecological perspective seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach. Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.

@@

The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021; REF] (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.

In this approach we include spatial structure how an implicit approach where different areas are treated as separate fleets [REF]. Differences among strata have been identified from various perspectives of krill population dynamics [REF], as well as from the influence of environmental variability within Subarea 48.1 [REF]. This approach allows integrate spatial heterogeneity into the stock assessment model to ensure that spatially structured processes are adequately represented. Although strata are treated as separate fleets, recruitment and growth processes remain unified, ensuring that the overall krill stock is modeled as a single population while accounting for spatially structured impacts at the Subarea 48.1 level. Each stratum is assigned a unique catch and survey history, reflecting differences in harvesting intensity and spatial exploitation patterns. Selectivity is modeled separately for each strata to account for potential variations in size composition due to fishing gear, depth distribution, or environmental conditions. Survey indices and length composition data are fleet-specific, allowing for variations in data collection methods and regional abundance estimates. The model considers interannual fluctuations of growth dynamics, maturity and reproduction, with a spawning pulse per year ocurring in January [@REF]. This method allows for spatial variability in exploitation patterns without explicitly modeling movement between areas. This approach enables a more realistic assessment of krill stock productivity under varying environmental conditions and harvesting scenarios, aligning with CCAMLR’s ecosystem-based management framework.

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others. Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including biomass index by strata, and biological characteristics of krill collected through net sampling. All this data was processed in @Dornam2021. Krill length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl. Mat. 2). For this analysis, Chlorophyll-a concentration (Chl) was assessed as a primary environmental driver. Chl data was downloaded from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and krill dynamics in the WAP. We applied a Generalized Linear Mixed Model (GLMM) to identify the environmental variable with main influence on krill population dynamics in the Antarctic Peninsula. Krill *Length* was used as a proxy for growth, incorporating both fixed and random effects to account for environmental and spatial variability. The icon model was formulated as:

$$
\text{Length}_i = \beta_0 + \beta_1 \text{ID}_i + \beta_2 \text{SIC}_i + \beta_3 \text{SST}_i + \beta_4 \text{Chla}_i + \beta_5 (\text{SST} \times \text{Chla})_i + (1 | \text{Year}_i) + \epsilon_i
$$

where $\beta_0$ is the intercept, and the fixed effects include *ID* (management unit or stratum), sea ice concentration (SIC), sea surface temperature (SST), and chlorophyll-a concentration (Chl). An interaction term between SST and Chl was included to capture potential synergistic effects on krill growth. A random effect for *Year* was added to account for interannual variability, mitigating biases from year-specific influences and data scarcity across strata. The residual error, $\epsilon_i$, was assumed to follow a normal distribution. The estimated coefficients and their 95% confidence intervals (CIs) were evaluated to determine the significance of each predictor, where coefficients overlapping zero indicated no significant relationship with krill length. Variance components were used to assess the proportion of variability explained by the random factors. Model selection was based on performance, with six different models tested, including one using the 75th percentile of lengths as an alternative response variable. The GLMM analysis was conducted using the *lme4* and *easystats* packages [@REF], while spatial data processing and stratification were performed with the *sf* and *CCAMLRGIS* packages [@REF]. Data handling for modeling purposes was carried out using the *tidyverse* library and its dependencies [@REF]. A full description of the models used is provided in Supplementary Material 2.

```{r eval=TRUE}
data <- data.frame(
  Source = c("Krill Fishery Monitoring Program",
                 "Krill Fishery Monitoring Program, collected by scientific observers under CCAMLR SISO program (vessel, nationality, georeferenced)",
                 "Krill Fishery Monitoring Program",
                 "AMLR Research Cruises Program", 
                  "Satellite data",
                 "Predator Abundance",
                 "AMLR Research Cruises Program"),
  Description = c("Catches",
                  "CPUE (standarized by vessel, area and deep)", 
                  "Krill Length compositions (grouped by 0.2 cm)",
                  "Biomass index from scientific AMLR cruises",
                  "Satellite-based chlorophyll-a concentration (monthly, interpolated) from the EU Copernicus Marine Service (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).",
                  "Data on penguin colonies, sourced from 'MAPPPD'",
                  "Predator Length composition krill diet"),
    TimeSeries = c("1997-2020",
                 "1997-2020",
                 "2000-2020",
                 "1991-2011 (based on the research cruise periods).",
                 "Monthly data from 1997 to present (based on satellite data availability).",
                 "Data from 2000s-present.",
                 "1977-2020 (from 2000 used in this analysis)"),
   Scale = c("by Strata",
                           "by Strata",
                           "by Strata",
                           "by Strata",
                           "48.1",
                           "by colony",
                           "by colony"))
ft <- flextable(data) %>%  
  set_caption("Comparison of Model Parameters") %>%  
  autofit() %>%  
  theme_vanilla() %>%  
  fontsize(size = 10, 
           part = "all") %>%  
  font(fontname = "Arial", 
       part = "all") %>%  
  set_table_properties(layout = "autofit", width = .75)
ft
```

@@

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a Catch-to-Length model with age dynamics was used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables. Regarding this, we consider a length-based model as an alternative to the age-based approach, especially given the scarcity (or null) of direct age data in krill monitoring. Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017]. In this sense, reading hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap. Length offers insights into the age structure of the population, as there is a correlation between age and length. So far, all integrated models for krill have been of the Catch-at-Length type [@Kinzey2015a; @Kinzey2019a; @Wang2021]. This means they are approaches that internally model age through growth parameters and the Age-length-key (ALK). To perform this conversion, various methods have been used, but the AKL matrix is perhaps the most important one [@ICCA2003]. ALK are generated by aging a sub-sample of a population and used to estimate the age distribution from larger length samples. ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases. Seasonal or multi-annual applications require careful justification, as using an ALK from a single period can introduce significant bias. Length-stratified sampling is essential, and the process of developing ALKs is labor-intensive, requiring optimal data collection to ensure accuracy  [@ICCA2003]. Formulae exist to estimate the number of age determinations and length measurements is calculated as;

$$
\frac{\text{Number at age for a length group}}{\text{Number of fish aged in that length group}}
$$

The ALK for the time period is raised to the length distribution for that time period:

$$
\text{Raised numbers at age by length group} = \text{Numbers at length} \times \text{Proportion at age for that length}
$$

If the ALK lacks data for certain length groups, data from those groups may be assigned to adjacent ones in the ALK. This process generates an age-length distribution by gear and time period, but caution is needed to avoid significant biases, especially for larger lengths with a broad age range. The numbers at age are summed across length groups, combining variances from ageing and length sampling to produce the age composition for the specified period. Numbers at age for all gears can then be calculated. Numbers at age for all gears can be calculated as:

$$
\sum N_a \times \left(\frac{W_{ct}}{W_{cs}}\right)
$$

where $\sum N_a$ is the sum of sampled numbers at age, W~ct~ is the total commercial catch weight, and W~cs~ is the
sampled commercial catch weight. Variance due to ageing of numbers at age for all gears can be calculated as:

$$
\sum \text{Var}_a
$$
where $\sum Var_a$  is the sum of variances due to ageing. Variance due to length sampling of numbers at age can be calculated as:

$$
\sum \text{Var}_l
$$
where $\sum Var_a$  is the sum of variances due to length sampling. The variances should be raised by:

$$
\frac{W_{ct}}{W_{cs}}
$$
where W~ct~ is the total commercial catch weight, and W~cs~ is the sampled commercial catch weight.

The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK. The majority of ALKs are based on growth parameters according to the most suitable formula for the species. In the case of krill, these parameters are derived from the von Bertalanffy growth curve. In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV. This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomasses with model estimates of vulnerable biomass.  Weight-at-length was assumed known and calculated using @Maschette2020 parametres. in krill, a AKL is exposed in Figure \@ref(fig:AKL)).

```{r AKL, eval=TRUE, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in SUbarea 48.1"}
knitr::include_graphics('Figs/AKL.png')
```

The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21. SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions. SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment, and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population. All analysis were executed in R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023]. Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit1). Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).

```{r parainit1, eval=TRUE}
dir1.1<-here("s1.1") 
start1 <- SS_readstarter(file = file.path(dir1.1,
                                          "starter.ss"),
                              verbose = FALSE)
dat1 <- SS_readdat(file = file.path(dir1.1, start1$datfile),
                        verbose = FALSE)
ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
                                    start1$ctlfil),
                        verbose = FALSE,
                        use_datlist = TRUE, 
                   datlist = dat1)
fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
                                                "forecast.ss"),
                              verbose = FALSE)
parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
 row.names( parbio)<-c("Nat M",
                       "Lmin", 
                       "Lmax",
                       "VonBert K",
                       "CV young",
                       "CV old", 
                       "Wt a", 
                       "Wt b",
                       "L50%", 
                       "Mat slope")
 SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
 Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
 Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]

 parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
parInit %>% 
  kable(format = "latex",
        booktabs = TRUE, 
        caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_styling(latex_options = c("striped"), 
                full_width = FALSE, 
                font_size = 9) %>%
  pack_rows(index = c("Natural Mortality" = 1,
                      "Growth" = 5,
                      "Length-Weight Relation" = 2,
                      "Maturity" = 2,
                      "Stock-Recruit Relation" = 5,
                      "Catchability" = 2,
                      "Selectivity" = 4))
```

### Growth conditioning

Growth was modeled following a von Bertalanffy growth function with size-at-age parameterization. The model assumed linear growth below a predefined settlement age of 1.5 years. The asymptotic length ($L_{\infty}$) was determined by the estimated length-at-age, indicating that the model used $L_{\infty}$ as an upper growth boundary. Exponential decay of growth beyond maximum age was disabled, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints. Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.

@@

Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years. Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$. The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan. The model used a direct parameterization without additional offsets between sexes. Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters.

@@

Selectivity functions were modeled using a length-based logistic formulation for the fishery, survey, and predator fleets. The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width). The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width. The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm. Parameter estimation phases were assigned, with negative values indicating fixed parameters. The survey selectivity was also modeled using a length-based logistic function. Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm. The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values). These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets. The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm. assuming that the predator can capture a wider range of krill sizes. These parameters were estimated in later phases (positive phase values), allowing flexibility in their estimation. Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization.

@@

The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies. This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@REF]. The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value of 23, with a prior of 10. The parameter was freely estimated in phase 1.  The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability. The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts. Recruitment deviations were estimated to account for interannual variability in recruitment dynamics. The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term. The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years. Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed. Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment. The maximum bias adjustment factor was 0.8274. To prevent overparameterization, recruitment deviations were constrained between -5 and 5. Cyclic patterns in recruitment were assumed each tree and five years [@REF], and advanced recruitment options were enabled for model flexibility.

@@

## Scenarios
